% Copyright 2007-2024 Joakim Nilsson
%
% This file is part of Combo Whist.
%
% Combo Whist is free software: you can redistribute it and/or modify
% it under the terms of the GNU General Public License as published by
% the Free Software Foundation, either version 3 of the License, or
% (at your option) any later version.
%
% Combo Whist is distributed in the hope that it will be useful,
% but WITHOUT ANY WARRANTY; without even the implied warranty of
% MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
% GNU General Public License for more details.
%
% You should have received a copy of the GNU General Public License
% along with Combo Whist.  If not, see <http://www.gnu.org/licenses/>.

\documentclass[a4paper]{article} % Document class
\input{../language.tex}          % Language-dependent code
\input{../../rules.tex}          % Common code to all languages

% Title
\defTitle{Combo Whist}{The Taintless Card Game}

% Links text
\defLinks{For the latest version of the rules, visit \rulesUrl, or subscribe to \rssUrl.}

% Author
\author{By Joakim Nilsson}

% Date and version
\ifdefined\varDev
	\date{Development version (based on version \varVersion-\varLanguage)---\today}
\else
	\date{Version \varVersion-\varLanguage---\today}
\fi

% Document
\begin{document}
	%=============
	% Intro pages
	%=============

	\introPages
	\pagebreak

	%======
	% Body
	%======

	\section{Overview}
	Combo Whist is a trick-taking card game and a variant of Whist. It is easy to learn in a few minutes, but difficult to master.

	Combo Whist presents the players with an exceptionally large plethora of ways of play by having them negotiate a combination of rules to use before each hand. This is achieved by presenting the players with a \emph{reasonably small number of bids} that can be combined in a \emph{large number of ways} in order to make ``combo bids,'' which determine the rules. This is done in hope that the vast possibilities of controlling the rules in sophisticated ways favors skill in contrast to chance more than in other Whist variants, without making the game overly complicated---making the game fun to play for both casual and professional players.

	\paragraph{Number of players:}
	4 is preferred, but 3 to $\infty$ is also possible, with some adjustments to the rules.

	\paragraph{Requirements:}
	Standard 52-card deck, pen and paper. (The helper cards are \emph{optional} and can aid in organizing a player's thoughts, but have no real effect on the game.)

	\paragraph{Card rank:}
	From highest to lowest: A, K, Q, J, 10, 9, 8, 7, 6, 5, 4, 3, 2

	\section{How to play}
	\subsection{Preparations}
	If there are only 3 players, remove $\clubsuit 7$ as well as all the 8s, 9s and 10s from the deck, making 4 suits of consisting of 10, 10, 10 and 9 cards, respectively.

	Make a column for each player on the paper. This is done in order to keep track of each player's score and some additional information. When the paper has been prepared, choose a dealer at random.

	\subsection{Deal}
	A deal consists of two main parts---the \emph{bidding} and the \emph{hand}. Because it should be easier to understand the bidding if the hand is already understood, the hand is described first.

	A deal begins with the dealer dealing 13 cards to each player, whereafter the bidding starts. After the bidding is complete, the hand starts.

	After the hand, players' new scores are recorded, and a new deal begins with the next dealer being the player on the current dealer's left.

	\subsubsection{Hand}
	The hand plays similarly to how a hand is played in most Whist variants: The player to the \emph{declarer's}\footnote{The term ``declarer'' is explained in Section~\sectionref{bidding}.} right is called \emph{forehand} and leads by playing a card. Next, in a clockwise manner, the other players play one card each. The other players must---if they can---play cards which follow the suit of forehand's card. Otherwise they must discard any card or play a \emph{trump}\footnote{A trump is a card from the trump suit. The trump suit is determined during bidding.}. Unlike in some Whist variants, there is no obligation for a player to play a trump in case they are out of cards in the leading suit.

	The player who played the highest card in the leading suit brings home the \emph{trick}\footnote{That is, takes all the played and discarded cards and puts them face-down on the table.} unless a trump is played, in which case the player who played the highest trump takes the trick. The player who brought home the last trick leads the next one.

	\subsubsection{Bidding}
	\label{sec:bidding}
	In Combo Whist, players make \emph{combo bids}. A combo bid is composed of exactly one standard bid and any number (including zero) of unique special bids. The bids have specific rules associated with them, which are applied during the hand and during scoring.

	The players take turns by bidding in a clockwise manner and the player to the dealer's left makes the first bid. A player can either make a combo bid that ranks higher than the previous combo bid, or pass. A player who passes is out of the bidding and may therefore not make any new bids until the next deal. If all players pass, a \emph{redeal} occurs, where the same dealer deals again.

	A combo bid's rank is defined as the combined rank of all its constituent standard and special bids. A combo bid must have a rank of at least 1.

	A proposed time limit between bids is 30 seconds, with an extra 60 seconds before the first bid. However for beginners, higher limits or no limits are recommended. A player who does not make a bid within the given time limit, passes automatically. The bidding continues until all players but one have passed. That non-passing player is appointed declarer and the hand begins.

	The standard and special bids that are available to combine into a combo bid are listed in the tables in Section~\sectionref{standardBids}, and Section~\sectionref{specialBids}, respectively. The number of tricks to bring home in order to complete a combo bid is listed in the standard bids table in the ``Tricks'' column. A special bid cannot be combined with bids listed in its ``Incompatibility'' column.

	If it is unclear when an event triggered by a bid is supposed to occur, refer to the special bids' ``Priority'' column. The priority number listed in said column decides in what order the events specified by a bid's rules will take place. The bids with the lowest priority numbers go first. All standard bids have the priority number 0.

	\subsubsection{Scoring}
	After a hand has been played, the declarer scores a number of points determined by what combo bid was bid and whether it was completed. If the bid was completed, they score as many points as as listed in the ``Score'' column for the standard bid. If the bid was not completed, the declarer loses 2 points. A player is allowed to attain a negative score, but if a player has a score below $-5$, they are not allowed to take part in the bidding. However, said players automatically score 1 free point after every deal they participate in (even in the event that no one bids and a redeal occurs).

	\subsubsection{Winning}
	\label{sec:winning}
	There are two variants for determining the winner in Combo Whist: \emph{classic} and \emph{limited}.

	\paragraph{Classic:}
	The winner is the player who first attains or exceeds the \emph{winning score}. The winning score starts at 13, but 1 is subtracted from it each time all players have dealt one deal each---after one \emph{round}. The winning score decreases \emph{after} the score for the final hand has been recorded. This continues until the winning score reaches 1, where it stays until someone wins. A player must win by completing a bid and can therefore not win merely because the winning score decreases. A player can also not win unless they have the solitary highest score.

	\paragraph{Limited:}
	A predetermined number of rounds are first played (one suggestion is 3). After all rounds have been played, the next player wins who completes a bid which results in said player attaining the solitary highest score. To clarify: A player can not win at the end of the last of the predetermined number of rounds.

	\paragraph{Win of Shame:}
	Common to both variants is the following rule: If all players but one attains a score of $-5$ or lower, the player with the highest score wins, regardless of the winning score. This type of win is called a \emph{Win of Shame}.

	\section{Miscellaneous}
	\subsection{Rules for more than 4 players}
	If there are more than 4 players participating in the game, for each deal, all players but 4 \emph{sit out}; That is, they don't participate in the deal. These are the players seated closest to the right of the dealer.
		
	\subsection{Talking}
	Players are allowed to talk on the condition that they don't hint about what cards they have.
		
	\subsection{Cheating}

	If the declarer unintentionally cheats, the current combo bid is not completed. If a non-declarer unintentionally cheats, the following occurs: A number of points equal to the current bid's score is subtracted from the cheating player's score. The current combo bid continues, but no subtraction of points is done from the declarer's score should the bid not be completed.

	If unintentional cheating occurs before a declarer has been appointed, the deal is canceled, and 2 points are subtracted from the cheating player's score.

	However, if all of the deal's players agree about how the events after the cheating occurred can be reverted, they should be reverted in the agreed-upon manner, without other changes to the scores except that 1 point is subtracted from the cheating player's score.

	A player who intentionally cheats in Combo Whist is never again allowed to play it because it is obvious that the do not respect the game's magnificence.

	%============
	% Bid tables
	%============

	\pagebreak
	\newgeometry{left=1cm, right=1cm, top=1cm}

	\section{Standard bids}
	\label{sec:standardBids}
	\begin{center}
		\begin{tabularx}{\textwidth}{
				p{2.5cm}
				S
				S
				ccX
			}

			\textbf{D\scriptsize ESIGNATION} &
			\rotccw{\textbf{Rank}} &
			\rotccw{\textbf{Score}} &
			\rotccw{\textbf{Trump}} &
			\rotccw{\textbf{Tricks}} &
			\textbf{Rules}
			\\[-3ex]

			\directlua{tableItemsStandardBids()}
		\end{tabularx}
	\end{center}

	\section{Special bids}
	\label{sec:specialBids}
	\begin{center}
		\begin{tabularx}{\textwidth}{
				p{2.3cm}
				S
				S
				p{2.7cm}
				X
			}

			\textbf{D\scriptsize ESIGNATION} &
			\rotccw{\textbf{Rank}} &
			\rotccw{\textbf{Priority}} &
			\textbf{Incompatibility} &
			\textbf{Rules}
			\\[-3ex]

			\directlua{tableItemsSpecialBids()}
		\end{tabularx}
	\end{center}
\end{document}
